The Design, Semantics, and Implementation of \CEU: a Synchronous Reactive Language based on Esterel

### 2015-10-09 paper.tex
Revisao Terra

Terra is a system for programming wireless sensor network applications that 
uses \CEU as its scripting language~\cite{ceu.terra}.
%
Figure \ref{fig.terra} shows the three basic elements of Terra:
\CEU as the scripting language,
a set of customized pre-built components,
and the embedded virtual-machine engine which can disseminate and install 
bytecode images dynamically.
%
This approach aims to combine the flexibility of remotely uploading code with 
the expressiveness and safety guarantees of \CEU.

@@

@@

\begin{figure}
\begin{minipage}[t]{0.50\linewidth}
\begin{lstlisting}[morekeywords={function}]
// Output events
output void REQUEST_TEMPERATURE;
output int  REQUEST_SEND;    // sends int

// Input events
input int  TEMPERATURE_DONE; // receives int
input void SEND_DONE;

// System calls
function int getRadioID (void);
\end{lstlisting}
\end{minipage}
%
\begin{minipage}[t]{0.50\linewidth}
\begin{lstlisting}[numbers=left,xleftmargin=2.5em]
// Output events
void VM.out(int evt_id, void* args) {
    switch (id){
        case O_REQUEST_TEMPERATURE:
            call TINYOS_TEMP.read();
        <...>; // O_REQUEST_SEND
    }
}

@@

// System calls
void VM.function(int id, void* params) {
    switch (id) {
        case F_GET_RADIO_ID:
            VM.push(TINYOS_NODE_ID);
    }
}
\end{lstlisting}
\end{minipage}
%
\caption{
%{\small %\textmd{
\CEU interface (in the left) with associated customized \VM (in the right).
%
\code{VM.out} redirects all output events to the corresponding OS call (ln.  
1--8).
%
Each \emph{TinyOS} event callback calls \code{VM.enqueue} for the corresponding 
input event (ln 10--14).
%
System calls use \code{VM.push} for immediate return values (ln. 16--22).
%
\label{lst.terra.defs}
}
\end{figure}

Considering the compilation process illustrated in Figure~\ref{fig.impl} for 
the \C back end, the main difference resides in the \emph{Code Generation} 
phase, which here outputs assembly instructions for the Terra \VM, instead of 
statements in \C.
%
The \VM includes special instructions to support complex and recurrent 
operations from the runtime of \CEU, such as dispatching events, clearing 
aborted trails,
TODO
\begin{comment}
Basically, they map

@@

A number of virtual machines have been proposed for embedded systems.
%
The \emph{Sun SPOT} Java-based platform with the \emph{Squawk JVM} brings Java 
for the embedded domain~\cite{wsn.sunspot}, but requires a much powerful 
hardware.%
\footnote{The \emph{Sun SPOT} uses a 32-bit CPU with 4 Mbytes of FLASH and 512 
KBytes of SRAM.}
%
\emph{Darjeeling}~\cite{wsn.darjeeling} and \emph{TakaTuka}~\cite{wsn.takatuka} 
are complete \emph{JVMs} targeting constrained embedded systems with support 
for multithreading and garbage collection.
%
Java has antagonistic design choices in comparison to \CEU:
it does not impose static bounds on memory usage and execution time, and 
provides preemptive multithreading with synchronization primitives for 
accessing shared memory.
%
Plummer et al.~\cite{esterel.vm} propose a Esterel-based \VM with similar 
design choices to our work.
To reduce code size, the \VM has a specialized instruction set to deal with 
events and concurrency constructs that are particular to Esterel.
However, the proposed \VM is only a proof of concept, with no support for 
arithmetic operations, external system calls, or remote reprogramming.

>>>>

Terra is a system for programming wireless sensor network applications which 
uses \CEU as its scripting language~\cite{ceu.terra}.
%
Figure \ref{fig.terra} shows the three basic elements of Terra:
\CEU as the scripting language,
a set of customized pre-built components,
and the embedded virtual-machine engine which can disseminate and install 
bytecode images dynamically.
%
This approach aims to combine the flexibility of remotely uploading code with 
the expressiveness and safety guarantees of \CEU.

@@

The main difference between the standard \C back end and the Terra \VM is the 
\emph{Code Generation} phase (as illustrated in Figure~\ref{fig.impl}), which 
here outputs assembly instructions for the \VM, instead of statements in \C.
%
To reduce the memory footprint of applications, the \VM includes special 
instructions for complex and recurrent operations from the runtime of \CEU, 
such as handling events and trails.

@@

\begin{figure}
\begin{minipage}[t]{0.51\linewidth}
\begin{lstlisting}[morekeywords={function}]
// Output events
output void REQUEST_TEMPERATURE;
output int  REQUEST_SEND;    // sends int value

// Input events
input int  TEMPERATURE_DONE; // recvs int value
input void SEND_DONE;

// System calls
function int getRadioID (void);
\end{lstlisting}
\end{minipage}
%
\begin{minipage}[t]{0.49\linewidth}
\begin{lstlisting}[numbers=left,xleftmargin=2.5em]
// Output events
void VM.out(int evt_id, void* args) {
    switch (id){
        case O_REQUEST_TEMPERATURE:
            call TINYOS_TEMP.read();
        <...>; // O_REQUEST_SEND
    }
}

@@

// System calls
void VM.function(int id, void* params) {
    switch (id) {
        case F_GET_RADIO_ID:
            VM.push(TINYOS_NODE_ID);
    }
}
\end{lstlisting}
\end{minipage}
%
\caption{
%{\small %\textmd{
\CEU interface (in the left) with customized \VM (in the right).
%
The routine \code{VM.out} redirects all output events to the corresponding OS 
calls (ln.  1--8).
%
Each \emph{TinyOS} event callback calls \code{VM.enqueue} for the corresponding 
input event (ln 10--14).
%
System calls use \code{VM.push} for immediate return values (ln. 16--22).
%
\label{lst.terra.defs}
}
\end{figure}

\begin{comment}
Basically, they map

@@

A number of virtual machines have been proposed for embedded systems.
%
The \emph{Sun SPOT} platform with the \emph{Squawk JVM} brings Java for the 
embedded domain~\cite{wsn.sunspot}, but requires a much powerful hardware.%
\footnote{The \emph{Sun SPOT} uses a 32-bit CPU with 4 Mbytes of FLASH and 512 
KBytes of SRAM.}
%
\emph{Darjeeling}~\cite{wsn.darjeeling} and \emph{TakaTuka}~\cite{wsn.takatuka} 
are complete \emph{JVMs} targeting constrained embedded systems with support 
for multithreading and garbage collection.
%
Java has antagonistic design choices in comparison to \CEU:
it does not impose static bounds on memory usage and execution time, and 
provides preemptive multithreading with synchronization primitives for 
accessing shared memory.
%
Plummer et al.~\cite{esterel.vm} propose a Esterel-based \VM with similar 
design choices to our work.
To reduce code size, the \VM has a specialized instruction set to deal with 
events and concurrency constructs that are particular to Esterel.
However, the proposed \VM is only a proof of concept, with no support for 
arithmetic operations, external system calls, or remote reprogramming.
